---
title: "Practical Machine Learning - Prediction Assignment Writeup Project"
author: "Nicholas Ong"
date: "11/18/2020"
output: html_document
---
# Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

# The Data
The training data for this project are available here: https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here: https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har

# Data Preparation
```{r, message=FALSE, warning=FALSE}
# Load required libraries
library(caret)
library(corrplot)
library(dplyr)
library(e1071)
library(ggplot2)
library(kernlab)
library(knitr)
library(lattice)
library(plyr)
library(randomForest)
library(rattle)
library(rpart)
library(rpart.plot)
library(RColorBrewer)

# Read data files
train_df <- read.csv('pml-training.csv', na.strings = c('NA', '#DIV/0!', ''))
test_df <- read.csv('pml-testing.csv', na.strings = c('NA', '#DIV/0!', ''))

dim(train_df)
dim(test_df)
```
We can see that there are 19,622 observations in the training dataset, vs. 20 observations in the testing dataset. There are 160 variables in both dataset. 

# Data Cleaning
Let us remove rows with missing data, and drop variables we are not interested in.
```{r}
train_df <- train_df[, colSums(is.na(train_df)) == 0]
test_df <- test_df[, colSums(is.na(test_df)) == 0]
train_df <- train_df[, -c(1:7)]
test_df <- test_df[, -c(1:7)]
```

# Model & Validation
We split the cleaned training dataset training-validation set with a ratio of 70%-30%, so as to compute out-of-sample errors/ cross validation. We will be applying 2 different models: (1) Random Forest, (2) SVM. Seed will be used for reproducability purposes.
```{r}
set.seed(128)
inTrain <- createDataPartition(train_df$classe, p = .7, list = F)
train <- train_df[inTrain, ]
valid <- train_df[-inTrain, ]
```

## Random Forest
Let's train and test our model using Random Forest.
```{r}
train$classe <- factor(train$classe)
model_RF <- randomForest(classe ~ ., data = train, method = 'class')
valid_RF <- predict(model_RF, valid, type = 'class')
confusionMatrix(valid_RF, as.factor(valid$classe))
```
We can see that estimated accuracy for Random Forest is 99.7% and estimated out-of-sample error is ~0.3%.

## SVM
Let's now train and test our model using SVM.
```{r}
model_SVM <- svm(classe ~ ., data = train)
valid_SVM <- predict(model_SVM, valid)
confusionMatrix(valid_SVM, as.factor(valid$classe))
```
We can see that estimated accuracy for Random Forest is only 94.6% and estimated out-of-sample error is ~5.4%.

# Conclusion
Given the accuracy results above, we can see that Random Forest is a better predicting model than SVM.

## Prediction
As such, below is the prediction results for 20 different test cases using Random Forest.
```{r}
predict_RF <- predict(model_RF, test_df, type = 'class')
predict_RF
```